---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# R/`delayed`

[![Travis-CI Build Status](https://travis-ci.org/jeremyrcoyle/delayed.svg?branch=master)](https://travis-ci.org/jeremyrcoyle/delayed)
[![Build status](https://ci.appveyor.com/api/projects/status/wid97j656ga0elme?svg=true)](https://ci.appveyor.com/project/jeremyrcoyle/delayed)
[![Coverage Status](https://img.shields.io/codecov/c/github/jeremyrcoyle/delayed/master.svg)](https://codecov.io/github/jeremyrcoyle/delayed?branch=master)
[![Project Status: WIP – Initial development is in progress, but there has not yet been a stable, usable release suitable for the public.](http://www.repostatus.org/badges/latest/wip.svg)](http://www.repostatus.org/#wip)
[![License: GPL v3](https://img.shields.io/badge/License-GPL%20v3-blue.svg)](http://www.gnu.org/licenses/gpl-3.0)

> A framework for parallelizing dependent tasks

__Author:__ [Jeremy Coyle](https://github.com/jeremyrcoyle)

---

## What's `delayed`?

`delayed` is an R package that provides a framework for parallelizing dependent
tasks in an efficient manner. It brings to R a subset of the
functionality implemented in Python's [Dask
library](https://dask.pydata.org/en/latest/). For details on how best to use
`delayed`, please consult the package
[documentation](https://nhejazi.github.io/delayed/) and
[vignette](https://nhejazi.github.io/delayed/articles/delayed.html)
online, or do so from within [R](https://www.r-project.org/).

---

## Installation

<!--
For standard use, we recommend installing the package from
[CRAN](https://cran.r-project.org/) via

```{r cran-installation, eval = FALSE}
install.packages("delayed")
```
-->

Install the most recent _stable release_ from GitHub via
[`devtools`](https://www.rstudio.com/products/rpackages/devtools/):

```{r gh-master-installation, eval = FALSE}
devtools::install_github("jeremyrcoyle/delayed")
```

---

## Issues

If you encounter any bugs or have any specific feature requests, please [file an
issue](https://github.com/jeremyrcoyle/delayed/issues).

---

## Example

This minimal example shows how to use `delayed` to handle dependent computations
via chaining of tasks:

```{r delay_fun_chain}
library(delayed)

# delay a function that does a bit of math
mapfun <- function(x, y) {(x + y) / (x - y)}
delayed_mapfun <- delayed_fun(mapfun)

set.seed(14765)
library(future)
plan(multicore, workers = 2)
const <- 7

# re-define the delayed object from above
delayed_norm <- delayed(rnorm(n = const))
delayed_pois <- delayed(rpois(n = const, lambda = const))
chained_norm_pois <- delayed_mapfun(delayed_norm, delayed_pois)

# compute it using the future plan (multicore with 2 cores)
chained_norm_pois$compute(nworkers = 2, verbose = TRUE)
```

_Remark:_ In the above, the delayed computation is carried out in parallel using
the framework offered by the excellent [`future`
package](https://github.com/HenrikBengtsson/future).

---

## License

&copy; 2017 [Jeremy R. Coyle](https://github.com/jeremyrcoyle)

The contents of this repository are distributed under the GPL-3 license. See
file `LICENSE` for details.

